Project 4

### 2018-11-21 Project4/report/Project4.tex
Finished

\begin{abstract}
In this study of phase transformations of a magnetic system we will estimate the critical temperature using the Ising model. To study the system of time we will use the Metropolis algorithm, which is a Markov chain-Monte Carlo method. Since we do not know all the possible states of the system beforehand, the Metropolis algorithm is beneficial to use, since it eliminates the partition function. The study will show that for a cold system of temperature 1 kT/J the system rarely has states of higher energies, when increasing the temperature to 2.4 kT/J the total energy of the system is increased, and we see that the probability for the system to be in a higher energy state is increased. When we look at how the system evolves as a function of temperature we clearly see that there is a phase transformation at a temperature close to the critical temperature. When estimating the critical temperature we find this to be $T_C = 2.256$ kT/J, which is fairly close to the analytical value found by Lars Onsager \cite{bib:onsager}.

@@

\subsection{Implementation}
For a MC cycle we initialize the lattice by setting all the spins in either a random or an ordered state, and then calculating the energy and the magnetic moment. We then iterate over the lattice at random positions (i.e. we check $L^2$ random spins), for each of these positions we calculate the change in energy $\Delta E$. Consider the ratio of the probability of the new energy $P\left(E_{new}\right)$ and the previous energy $P\left(E_{prev}\right)$:
%
\begin{align*}
    \frac{P\left(E_{new}\right)}{P\left(E_{prev}\right)} = \frac{e^{-\beta E_{new}}/Z}{e^{-\beta E_{prev}}/Z} = e^{-\beta\left(E_{new} - E_{prev}\right)} = e^{-\beta\Delta E},
\end{align*}
%
where Z is the the partition function. We now perform the Metropolis test for a given $\Delta E$;
%
\begin{equation*}
    r \leq e^{-\beta \Delta E}, \qquad r \in [0,1]
\end{equation*}
%
i.e. if the ratio is larger than or equal to a random number $r \in [0,1]$ we accept a new configuration, and flip the spin.  We are looking at the change in energy if we were to flip a random spin; if the change of energy is large, the likelihood of this spin being flipped is small, so we reject this solution, on the other hand, if the probability is large, we accept the configuration. Since we expect the system to move to more likely state, we are looking for states that does not increase the systems energy. A large change in energy means we are looking at a spin that has half or more of its surrounding spins in the same configuration, and the system seems to locally move towards a steady state, therefore the local system is not changed. For each such Metropolis test that succeeds we add the flip of spin as a contribution to the systems energy and magnetic moment. When the calculations over the lattice is complete we conclude a single MC cycle and store the final value for the energy and magnetic moment.

@@

\begin{thebibliography}{}
\bibitem{bib:onsager} Onsager, Lars,
 \textit{Crystal Statistics. I. A Two-Dimensional Model with an Order-Disorder Transition}
\url{https://link.aps.org/doi/10.1103/PhysRev.65.117}
\end{thebibliography}

>>>>

\begin{abstract}
In this study of phase transformations of a magnetic system we will estimate the critical temperature using the Ising model. To study the system of time we will use the Metropolis algorithm, which is a Markov chain-Monte Carlo method. Since we do not know all the possible states of the system beforehand, the Metropolis algorithm is beneficial to use, since it eliminates the partition function. The study will show that for a cold system of temperature 1 kT/J the system rarely has states of higher energies, when increasing the temperature to 2.4 kT/J the total energy of the system is increased, and we see that the probability for the system to be in a higher energy state is increased. When we look at how the system evolve as a function of temperature we clearly see that there is a phase transformation at a temperature close to the critical temperature. When estimating the critical temperature we find this to be $T_C = 2.256$ kT/J, which is fairly close to the analytical value found by Lars Onsager \cite{bib:onsager}.

@@

\subsection{Implementation}
For a MC cycle we initialize the lattice by setting all the spins in either a random or an ordered state, and then calculating the energy and the magnetic moment. We then iterate over the lattice at random positions (i.e. we check $L^2$ random spins), for each of these positions we calculate the change in energy $\Delta E$. Consider the ratio of the probability of the new energy $P\left(E_{new}\right)$ and the previous energy $P\left(E_{prev}\right)$:
%
\begin{align*}
    \frac{P\left(E_{new}\right)}{P\left(E_{prev}\right)} = \frac{e^{-\beta E_{new}}/Z}{e^{-\beta E_{prev}}/Z} = e^{-\beta\left(E_{new} - E_{prev}\right)} = e^{-\beta\Delta E},
\end{align*}
%
where Z is the the partition function. We now perform the Metropolis test for a given $\Delta E$;
%
\begin{equation*}
    r \leq e^{-\beta \Delta E}, \qquad r \in [0,1]
\end{equation*}
%
i.e. if the ratio is larger than or equal to a random number $r \in [0,1]$ we accept a new configuration, and flip the spin.  We are looking at the change in energy if we were to flip a random spin; if the change of energy is large, the likelihood of this spin being flipped is small, so we reject this solution, on the other hand, if the probability is large, we accept the configuration. Since we expect the system to move to more likely state, we are looking for states that does not increase the systems energy. A large change in energy means we are looking at a spin that has half or more of its surrounding spins in the same configuration, and the system seems to locally move towards a steady state, therefore the local system is not changed. For each such Metropolis test that succeeds we add the flip of spin as a contribution to the systems energy and magnetic moment. When the calculations over the lattice is complete we conclude a single MC cycle and store the final value for the energy and magnetic moment. When we are calculating the final mean values of the energy and magnetic moment, we ignore the first 5\% of the data, so we are not including the values before the system is in equilibrium.

@@

\begin{thebibliography}{}
\bibitem{bib:onsager} 
Onsager, Lars,
 \textit{Crystal Statistics. I. A Two-Dimensional Model with an Order-Disorder Transition}
\url{https://link.aps.org/doi/10.1103/PhysRev.65.117}

\end{thebibliography}
